Sperm should evolve to make female meiosis fair.

### 2014-07-25 writeup_and_notes/GeneticsResub/reviews.tex

\begin{singlespace}
	{\bf{Response -- }}  We believe that this is an important phenomenon. 
	We have modified the manuscript to clarify that it is  quite impressive that we find evidence
	for male influence on female drive in BOTH cases where this phenomenon could possibly be observed.
	We note that although direct evidence for female meiotic drive is modest, 
	most experiments to dates have lacked the power to detect most modest drivers. 
	We believe that given our theoretical predictions and the affordability of large scale GBS experiments,
	data weighing in on this paper is on its way. 
	We appreciate the opportunity to get ahead of this data by offering strong predictions that could guide analyses.
\end{singlespace}
\end{enumerate}

@@

\begin{enumerate}
\item	The importance of the wok hinges critically on the the degree to which:  {\bf{\emph{i)}}}
female meiotic drive operates in nature, and {\bf{\emph{ii)}}} sperm influence the level of
female meiotic drive.  In species like maize, where every chromosome has two or
more established meiotic drivers (``chromosomal knobs''), the importance of
meiotic drive is obvious, but in most species there is little or no evidence for
the current operation of meiotic drive -- although low levels of transmission
ratio distortion (TRD) are becoming increasingly well established from
genome-wide SNP studies.  However, the rapid divergence of telomeric sequences
among closely related species is certainly consistent with the widespread
operation of transient episodes of meiotic drive via centromere drive.  In the
present manuscript, the prevalence and importance of female meiotic drive is
quickly summarized by the terse statements ``Meiotic drivers observed in both
plants [5, 10, 9], and animals [1, 29, 21] highlight this conflict -- the selfish
benefits and the associated pleiotropic fitness costs of drive sustain a
balanced polymorphism [24], and often generate ongoing evolutionary escalations
of drive suppressors and enhancers [8, 9].''  I think the paper would be far more
convincing and appreciated if the authors put a tad more effort into convincing
the reader that female meiotic drive is sufficiently common to be a widespread
selective force.

\item	 To my knowledge, the evidence for sperm influencing the outcome of meiotic
drive is restricted to the In and Om drive systems in mice and aster transfer by
the sperm in C. elegans.  These are clearly proof of process examples but I
think that many readers will take issue with the assumed general ability of
sperm to substantially influence the outcome of female meiotic drive. This
skepticism is to be expected  because the egg typically completes meiosis
immediately after fertilization, at a time when the sperm pronucleus is
decondensing and presumably not yet transcriptionally active. The more evidence
that the authors can provide for sperm being capable of influencing female
meiotic drive, the better this paper will be received. The authors could bolster
their case by briefly reviewing the recent evidence for extensive paternal
effects via RNA and protein packaged in the sperm, delivered to the egg, and
influencing subsequent early embryonic development.

\item	Finally, female meiotic drive potentially can be suppressed by alleles
expressed in the female (traditional models) or by alleles originating in the
sperm and acting in the egg immediately after fertilization (the authors new
model).  It would be useful to compare the constraints for the evolution of both
types of suppressors to evaluate if sperm-mediated suppression of female meiosis
has some evolutionary advantage that make it easier to evolve or non-redundant
to female-coded suppressors.
\end{enumerate}

\subsection*{Minor Points:}
\begin{enumerate}
\item   On the top of page 7 the statement ``This analytical approximation, derived
from Equation (1) assuming Hardy-Weinberg'', -- precedes the introduction of
Equation (1) in the Appendix.  I suggest pointing out here that all equations
described in the text are located in the Appendix (so the reader doesn't stop
reading and start looking for them, thinking they somehow overlooked them in the
body of the manuscript).
\yb{We can relate to the confusion and frustration of struggling to find the relevant equation. We have therefore added the parenthetical comment --   ``\emph{We note that Equation 1 and all equations discussed in the main text are presented in the appendix and derived in File S2}'' -- to the end of this paragraph.}

\item There are 6 models described in the text but these are not described in any
detail until the Appendix.  The authors might consider placing a table in the
main body of the text that tersely describes the 6 models.?
\yb{We have added a Methods section, which introduces these six models, and makes our manuscript more in line with the GENETICS format.}

\item On the bottom of page 14 the authors write ``we predict that a lower
opportunity for female meiotic drive, e.g. an animal lineage with a long history
of high inbreeding or selfing, would often be accompanied by a relaxation of the
requirement of fertilization for the completion of female meiosis.''  As I
recall, species with a history if inbreeding or selfing are only at the tips of
most phylogenetic trees (i.e., they do not persist over geological time and do
not have a ``LONG history of high inbreeding or selfing'') so this prediction may
be limited.

\item On the bottom of page 14, do you also predict that observed drive in species
hybrids will depend on the direction (which male) is used in the backcross?

\item I think that equations like $(s < d -1/2 + 3sh/2 - dsh)$ need parentheses to be
clear. 
\yb{To add clarity, we have surrounded longer in-text equations like the example above, with parentheses.}

>>>>

\begin{singlespace}
	{\bf{Response -- }}  We believe that this is an important phenomenon. 
	We have modified the manuscript to clarify that it is  quite impressive that we find evidence
	for male influence on female drive in BOTH cases where this phenomenon could possibly be observed.
	We note that although direct evidence for female meiotic drive is modest, 
	most experiments to dates have lacked the power to detect most modest drivers. 
	We believe that given our theoretical predictions and the affordability of large scale GBS experiments,
	data weighing in on this paper are on its way. 
	We appreciate the opportunity to get ahead of this data by offering strong predictions that could guide analyses.
\end{singlespace}
\end{enumerate}

@@

\begin{enumerate}
\begin{singlespace}
\item	\emph{The importance of the wok hinges critically on the the degree to which:  {\bf{\emph{i)}}}
female meiotic drive operates in nature, and {\bf{\emph{ii)}}} sperm influence the level of
female meiotic drive.  In species like maize, where every chromosome has two or
more established meiotic drivers (``chromosomal knobs''), the importance of
meiotic drive is obvious, but in most species there is little or no evidence for
the current operation of meiotic drive -- although low levels of transmission
ratio distortion (TRD) are becoming increasingly well established from
genome-wide SNP studies.  However, the rapid divergence of telomeric sequences
among closely related species is certainly consistent with the widespread
operation of transient episodes of meiotic drive via centromere drive.  In the
present manuscript, the prevalence and importance of female meiotic drive is
quickly summarized by the terse statements ``Meiotic drivers observed in both
plants [5, 10, 9], and animals [1, 29, 21] highlight this conflict -- the selfish
benefits and the associated pleiotropic fitness costs of drive sustain a
balanced polymorphism [24], and often generate ongoing evolutionary escalations
of drive suppressors and enhancers [8, 9].''  I think the paper would be far more
convincing and appreciated if the authors put a tad more effort into convincing
the reader that female meiotic drive is sufficiently common to be a widespread
selective force.}
\end{singlespace}
\begin{singlespace}
	{\bf{Response -- }}  We agree and have added more support of the likely importance and prevalence of female meiotic drive. 
	We hope that new theoretical work (including ours) coupled with the falling expense of GBS, will encourage the discovery of additional drive systems in non-model organisms in the coming years.
\end{singlespace}

\begin{singlespace}
\item	 \emph{To my knowledge, the evidence for sperm influencing the outcome of meiotic
drive is restricted to the In and Om drive systems in mice and aster transfer by
the sperm in \emph{C. elegans}.  These are clearly proof of process examples but I
think that many readers will take issue with the assumed general ability of
sperm to substantially influence the outcome of female meiotic drive. This
skepticism is to be expected  because the egg typically completes meiosis
immediately after fertilization, at a time when the sperm pronucleus is
decondensing and presumably not yet transcriptionally active. The more evidence
that the authors can provide for sperm being capable of influencing female
meiotic drive, the better this paper will be received. The authors could bolster
their case by briefly reviewing the recent evidence for extensive paternal
effects via RNA and protein packaged in the sperm, delivered to the egg, and
influencing subsequent early embryonic development. }
\end{singlespace}
\begin{singlespace}
	{\bf{Response -- }}  
	We have added a discussion of these points to our introduction. 
	Specifically, we now discuss the influence of sperm on basic processes in the egg, 
	and highlight that although data are preliminary $In$ and $Om$ are the best cases of polymorphic female drivers in animals, 
	and therefore current support for our theory is actually quite high. 
\end{singlespace}

\begin{singlespace}
\item	 \emph{Finally, female meiotic drive potentially can be suppressed by alleles
expressed in the female (traditional models) or by alleles originating in the
sperm and acting in the egg immediately after fertilization (the authors new
model).  It would be useful to compare the constraints for the evolution of both
types of suppressors to evaluate if sperm-mediated suppression of female meiosis
has some evolutionary advantage that make it easier to evolve or non-redundant
to female-coded suppressors. }
\end{singlespace}
\begin{singlespace}
	{\bf{Response -- }}  For comparison, we have added a female meiotic drive suppressor unlinked to a driver as model 
	$6^\prime$, and compared its evolution to a sperm acting driver.
	In general, the spread of an unlinked female drive suppressor is qualitatively similar to that of an unlinked sperm-acting drive suppressor; 
	however, the initial spread of the sperm suppressor is equal to (when the female suppressor is dominant), or more rapid than (otherwise) the female suppressor. 
	Given the difference in mechanism of female drive suppression in sperm as compared to females, sperm suppression of female drive is likely not functionally redundant with female drive suppression and provides a qualitatively different mechanism of suppression.
\end{singlespace}

\end{enumerate}

\subsection*{Minor Points:}
\begin{enumerate}
\begin{singlespace}
\item \emph{  On the top of page 7 the statement ``This analytical approximation, derived
from Equation (1) assuming Hardy-Weinberg'', -- precedes the introduction of
Equation (1) in the Appendix.  I suggest pointing out here that all equations
described in the text are located in the Appendix (so the reader doesn't stop
reading and start looking for them, thinking they somehow overlooked them in the
body of the manuscript). \\ \\}
{\bf{Response -- }} We can relate to the confusion and frustration of struggling to find the relevant equation. We have therefore added the parenthetical comment --   ``\emph{We note that Equation 1 and all equations discussed in the main text are presented in the appendix and derived in File S2}'' -- to the end of this paragraph. 
We also believe that the METHODS section, which we present before the results in this new manuscript, will help prepare readers for this. 
\end{singlespace}

\begin{singlespace}
\item \emph{There are 6 models described in the text but these are not described in any
detail until the Appendix.  The authors might consider placing a table in the
main body of the text that tersely describes the 6 models? \\ \\}
{\bf{Response -- }} We have added a Methods section, which introduces these six models, and makes our manuscript more in line with the GENETICS format.
\end{singlespace}

\begin{singlespace}
\item \emph{On the bottom of page 14 the authors write ``we predict that a lower
opportunity for female meiotic drive, e.g. an animal lineage with a long history
of high inbreeding or selfing, would often be accompanied by a relaxation of the
requirement of fertilization for the completion of female meiosis.''  As I
recall, species with a history if inbreeding or selfing are only at the tips of
most phylogenetic trees (i.e., they do not persist over geological time and do
not have a ``LONG history of high inbreeding or selfing'') so this prediction may
be limited. \\ \\ }
{\bf{Response -- }} We agree that biology (e.g. extinction) might conspire to make this a challenging 
prediction to test, and we point this out in our new manuscript. 
However, we believe that this hypothesis follows from our results and therefore still propose it.
\end{singlespace}

\begin{singlespace}
\item \emph{On the bottom of page 14, do you also predict that observed drive in species
hybrids will depend on the direction (which male) is used in the backcross?\\ \\}
{\bf{Response -- }}  We have expanded and clarified our prediction regarding interspecific crosses in our new manuscript. 
\end{singlespace}

\begin{singlespace}
\item 
\emph{I think that equations like $(s < d -1/2 + 3s_h/2 - ds_h)$ need parentheses to be clear. \\ \\ }
	{\bf{Response -- }}  To add clarity, we have surrounded longer in-text equations like the example above, with parentheses.
\end{singlespace}
%\yb{To add clarity, we have surrounded longer in-text equations like the example above, with parentheses.}
